---
title: "Portfolio 2"
---

```{r packages}
library(tidyverse)
```

I am reading in data from the [PTSD Repository](https://ptsd-va.data.socrata.com/), which was created by the National Center for PTSD to help people understand what is currently known about PTSD treatment. It includes 583 studies. 

```{r read-dataset}
PTSD <- read_csv("data/PTSD.csv")
```

I want to filter the dataset to only inciude participants who are active duty military or veterans. I also want the treatment focus at the study level to be on PTSD. Here I created a new dataframe with those filters. 

```{r filter-military-status}
PTSD_military <- PTSD %>% 
  filter(`Military Status (Study level)` %in% c("Military", "Veteran"), `Treatment Focus (Study level)` == "PTSD") #filter dataset for active duty military and veterans, as well as studys focusing only on PTSD
```

Now, I want to organize the types of psychotherapy into 5 groups so I can compare more easily. The 5 groups are: EMDR, TF-CBT, TF-Other, Non-trauma-focused, and Mixed psychotherapy. I took these groupings from a resource I found on the same site as the data. They give suggestions on how you might group together specific standardized treatments into larger categories. [Table](https://ptsd-va.data.socrata.com/Reference/Psychotherapy-Progressively-Specific-Grouping/5vmb-i9mg/about_data)
EMDR is Eye Movement Desensitization and Reprocessing. TF-CBT is trauma-focused cognitive behavioral therapy, which includes exposure therapy and cognitive therapy. TF-Other is tramua-focused, but not cognitive behavioral therapy: some examples include Reconsolidation of Traumatic Memories (RTM), and Trauma counseling. Non-trauma-focused includes CBT without trauma focus, and other non-trauma-focused treatments. Lastly, Mixed psychotherapy includes a combination of multiple treatments, for example CBT + EMDR. 
Not all of the standardized treatments are in the list i am pulling from, so I am going to name those as "other". I mainly want to focus on comparing these 5 main groups. 

First, I will add a new variable, "category". The levels will be the 5 different groups plus other: EMDR, TF-CBT, TF-Other, Non-trauma-focused, Mixed psychotherapy, and other. 

```{r new-category-variable}
PTSD_military <- PTSD_military %>% 
  mutate(
    category = case_when(
      `Standardized Treatment Name` == "Eye Movement Desensitization and Reprocessing (EMDR)" ~ "EMDR",
      `Standardized Treatment Name` %in% c("Prolonged Exposure (PE)", "Narrative Exposure Therapy (NET)", "Exposure therapy", "Imaginal exposure", "Imaginal exposure (experimental paradigm) + placebo", "Imaginal exposure + imagery rescripting", "Implosive Therapy", "Virtual reality exposure therapy (VRET)", "Written Exposure Therapy (WET)", "Written narrative exposure", "Imagery Rehearsal Training (IRT)", "Cognitive Behavioral Conjoint Therapy (CBCT)", "Impact of Killing (IOK)", "Trauma-focused cognitive behavioral therapy (TF-CBT)", "Cognitive Processing Therapy (CPT)", "Cognitive Therapy for PTSD (CT-PTSD)") ~ "TF-CBT",
      `Standardized Treatment Name` %in% c("Brief Eclectic Psychotherapy (BEP)", "Modified Lifespan Integration", "Multimodular motion-assisted memory desensitization and
reconsolidation (3MDR)", "Observed and Experiential Integration - Switching component", "Reconsolidation of Traumatic Memories (RTM)", "Trauma counseling", "Accelerated Resolution Therapy (ART)") ~ "TF-Other",
      `Standardized Treatment Name` %in% c("Behavioral activation (BA)", "Cognitive behavioral therapy (CBT) without a trauma focus", "Cognitive Behavioral Therapy for Insomnia (CBT-I)", "Cognitive restructuring and imagery modification", "Metacognitive Therapy", "Stress Inoculation Training (SIT)", "Affect management therapy", "Coping skills treatment", "Interpersonal Psychotherapy (IPT)", "Memory specificity training", "Non-trauma-focused counseling", "Present Centered Therapy (PCT)", "Psychodynamic psychotherapy", "PTSD psychotherapy", "Resilience-oriented treatment", "Skills Training in Affect and Interpersonal Regulation (STAIR)", "Supportive counseling", "Dialectical Behavior Therapy") ~ "Non-trauma-focused", 
      `Standardized Treatment Name` %in% c("Cognitive behavioral therapy (CBT) + Eye Desensitization and Reprocessing (EMDR)", "Cognitive Behavioral Therapy for Insomnia (CBT-I) + Imagery Rehearsal Training (IRT)", "Dialectical Behavior Therapy (DBT) + trauma-focused cognitive behavioral therapy (TF-CBT)", "Eye Movement Desensitization and Reprocessing (EMDR) + cognitive behavioral therapy (CBT) without a trauma focus", "Eye Movement Desensitization and Reprocessing (EMDR) + progressive muscle relaxation (PMR) + group psychoeducation", "Narrative Exposure Therapy (NET) + Interpersonal Psychotherapy (IPT)", "Prolonged Exposure (PE) + Stress Inoculation Training (SIT)", "Skills Training in Affect and Interpersonal Regulation (STAIR) plus modified Prolonged Exposure (PE)") ~ "Mixed-psychotherapy"
      )
    )
```















