---
title: "Portfolio 3"
output: html_document
---

```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))
library(tidytuesdayR)
library(tidyverse)
```

```{r import-data, echo=FALSE}
tt <- tt_load("2026-02-17")
```

This TidyTuesday dataset contains long-term agricultural statistics from New Zealand. 

# Glimpse Data

```{r glimpse}
map(tt, glimpse)
```

# Wrangle

```{r Wrangle}
df <- tt$dataset %>%
  janitor::clean_names() #changes names to snake_case (not needed for this dataset, but trying to make it a part of my workflow)
glimpse(df)
```
```{r count}
df %>% count(measure, sort = TRUE)
```

# Visualize

This visualization shows a steady long-term rise and fall in New Zealand's sheep industry. Sheep populations, breeded ewes, and lamb numbers all increased steadily until peaking around the early 1980s, followed by a substantial decline over the following decades. The trends are all very similar, suggesting that the overall sheep population and reproduction rates changed together. 

```{r visualize}
selected_measures <- c(
  "Total Sheep",
  "Total Lambs Marked and/or Tailed",
  "Breeding Ewes 2 Tooth and Over put to Ram"
)

df_compare <- df %>%
  filter(measure %in% selected_measures)

ggplot(df_compare,
       aes(x = year_ended_june,
           y = value / 1e6, #change y-axis so its not in scientific notation, but in millions
           color = measure)) +
  geom_line() +
  labs(
    title = "Trends in New Zealand Sheep Population and Reproduction",
    x = "Year",
    y = "Count (millions)",
    color = NULL
  ) +
  theme_minimal()
```
