---
title: "Portfolio 4"
subtitle: "PTSD: Part 1"
---

### Goal
The goal of this portfolio piece is to practice data cleaning, filtering, and visualization in R using a real-world clinical research dataset. I use data from the [PTSD Repository](https://ptsd-va.data.socrata.com/) created by the National Center for PTSD, which summarizes 583 studies on PTSD treatment. In this project, I focus on studies involving active-duty service members and veterans and examine how PTSD-focused psychotherapy treatments are distributed across broader treatment categories (e.g., TF-CBT, EMDR). This project is intentionally iterative; I use it to refine my code, improve visualization clarity, and develop a repeatable workflow for categorizing messy treatment labels.

```{r packages, include=FALSE}
library(tidyverse)
```

First, I read in the dataset from the [PTSD Repository](https://ptsd-va.data.socrata.com/). 

```{r read-dataset}
PTSD <- read_csv("data/PTSD.csv", show_col_types = FALSE)
```

Next, I filter the dataset to include only studies where participants are active duty military or veterans, and where the treatment focus at the study level is PTSD. I store this filtered data in a new data frame.

```{r filter-military-status}
PTSD_military <- PTSD %>% 
  filter(
    `Military Status (Study level)` %in% c("Military", "Veteran"),
    `Treatment Focus (Study level)` == "PTSD"
  )
```

Now, I organize standardized psychotherapy treatments into five broader groups so that they are easier to compare. The five groups are EMDR, TF-CBT, TF-Other, Non-trauma-focused, and Mixed psychotherapy. I based these grouping on a resource provided on the same database as the dataset: [Psychotherapy Progressively Specific Grouping table](https://ptsd-va.data.socrata.com/Reference/Psychotherapy-Progressively-Specific-Grouping/5vmb-i9mg/about_data). 
- EMDR = Eye Movement Desensitization and Reprocessing
- TF-CBT = trauma-focused cognitive behavioral therapy (including exposure and cognitive therapies)
- TF-Other = trauma-focused therapies that are not TF-CBT (e.g., RTM, trauma counseling)
- Non-trauma-focused = CBT without trauma focus and other non-trauma-focused approaches
- Mixed psychotherapy = combinations of multiple psychotherapy approaches (e.g., CBT + EMDR)

Not all standardized treatment names in the dataset appear in the grouping list I used. Any treatments that do not match one of the five groups are labeled as Other.
To do this, I create a new variable called category with six levels: EMDR, TF-CBT, TF-Other, Non-trauma-focused, Mixed psychotherapy, and Other.

```{r new-category-variable}
PTSD_military <- PTSD_military %>% 
  mutate(
    category = case_when(
      `Standardized Treatment Name` == "Eye Movement Desensitization and Reprocessing (EMDR)" ~ "EMDR",
      
      `Standardized Treatment Name` %in% c(
        "Prolonged Exposure (PE)", "Narrative Exposure Therapy (NET)", "Exposure therapy",
        "Imaginal exposure", "Imaginal exposure (experimental paradigm) + placebo",
        "Imaginal exposure + imagery rescripting", "Implosive Therapy",
        "Virtual reality exposure therapy (VRET)", "Written Exposure Therapy (WET)",
        "Written narrative exposure", "Imagery Rehearsal Training (IRT)",
        "Cognitive Behavioral Conjoint Therapy (CBCT)", "Impact of Killing (IOK)",
        "Trauma-focused cognitive behavioral therapy (TF-CBT)",
        "Cognitive Processing Therapy (CPT)", "Cognitive Therapy for PTSD (CT-PTSD)"
      ) ~ "TF-CBT",
      
      `Standardized Treatment Name` %in% c(
        "Brief Eclectic Psychotherapy (BEP)", "Modified Lifespan Integration",
        "Multimodular motion-assisted memory desensitization and reconsolidation (3MDR)",
        "Observed and Experiential Integration - Switching component",
        "Reconsolidation of Traumatic Memories (RTM)", "Trauma counseling",
        "Accelerated Resolution Therapy (ART)"
      ) ~ "TF-Other",
      
      `Standardized Treatment Name` %in% c(
        "Behavioral activation (BA)",
        "Cognitive behavioral therapy (CBT) without a trauma focus",
        "Cognitive Behavioral Therapy for Insomnia (CBT-I)",
        "Cognitive restructuring and imagery modification", "Metacognitive Therapy",
        "Stress Inoculation Training (SIT)", "Affect management therapy",
        "Coping skills treatment", "Interpersonal Psychotherapy (IPT)",
        "Memory specificity training", "Non-trauma-focused counseling",
        "Present Centered Therapy (PCT)", "Psychodynamic psychotherapy",
        "PTSD psychotherapy", "Resilience-oriented treatment",
        "Skills Training in Affect and Interpersonal Regulation (STAIR)",
        "Supportive counseling", "Dialectical Behavior Therapy"
      ) ~ "Non-trauma-focused",
      
      `Standardized Treatment Name` %in% c(
        "Cognitive behavioral therapy (CBT) + Eye Desensitization and Reprocessing (EMDR)",
        "Cognitive Behavioral Therapy for Insomnia (CBT-I) + Imagery Rehearsal Training (IRT)",
        "Dialectical Behavior Therapy (DBT) + trauma-focused cognitive behavioral therapy (TF-CBT)",
        "Eye Movement Desensitization and Reprocessing (EMDR) + cognitive behavioral therapy (CBT) without a trauma focus",
        "Eye Movement Desensitization and Reprocessing (EMDR) + progressive muscle relaxation (PMR) + group psychoeducation",
        "Narrative Exposure Therapy (NET) + Interpersonal Psychotherapy (IPT)",
        "Prolonged Exposure (PE) + Stress Inoculation Training (SIT)",
        "Skills Training in Affect and Interpersonal Regulation (STAIR) plus modified Prolonged Exposure (PE)"
      ) ~ "Mixed psychotherapy",
      
      .default = "Other"
    )
  )
```

To visualize how frequently each category appears in the filtered dataset, I plot the number of studies per category.

```{r plot}
PTSD_military %>% 
  count(category) %>% 
  ggplot(aes(x = reorder(category, n), y = n)) +
  geom_col(fill = "purple") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Number of studies",
    title = "PTSD Treatment Categories",
    subtitle = "Military/Veteran Studies"
  ) +
  theme_classic()
```

After viewing the plot, a large portion of treatments fall into Other, which suggests that many treatment labels in this dataset do not match my initial grouping list. As a next step, I inspect which treatment names appear most often in the Other category and decide whether they should be reclassified into one of the five main psychotherapy categories.

```{r sort-other}
PTSD_military %>% 
  filter(category == "Other") %>% 
  count(`Treatment Name`, sort = TRUE)
```

Given this output, several "Other" treatment names appear multiple times (e.g., placebo, waitlist, TAU, PCT, and several medication or control conditions). I will have to do futher research to find out where these treatments belong. 

### Next steps (work in progress)
Next, I plan to recode common “Other” treatments into the existing categories when appropriate, separate psychotherapy vs medication/control conditions, and explore whether category frequencies differ by study characteristics (e.g., population type, setting, or study design).
