---
title: "Portfolio 5"
---

For this portfolio piece, I want to build on my previous analysis of the PTSD Repository dataset (portfolio 2). I created an initial workflow for filtering studies and categorizing psychotherapy treatments. While exploring the results, I discovered that a large portion of studies were categorized as “Other,” suggesting that my initial categorization did not fully capture the structure of the dataset.
In this follow-up analysis, I refine the workflow by separating psychotherapy from medication and control conditions, and improving the categorization process. 

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)

PTSD <- read_csv("data/PTSD.csv", show_col_types = FALSE)
```

```{r recreate-p2-dataset}
PTSD_military <- PTSD %>% 
  filter(
    `Military Status (Study level)` %in% c("Military", "Veteran"),
    `Treatment Focus (Study level)` == "PTSD"
  )
```

```{r og-categories}
PTSD_military <- PTSD_military %>% 
  mutate(
    category = case_when(
      `Standardized Treatment Name` == "Eye Movement Desensitization and Reprocessing (EMDR)" ~ "EMDR",

      `Standardized Treatment Name` %in% c(
        "Prolonged Exposure (PE)", "Narrative Exposure Therapy (NET)", "Exposure therapy",
        "Imaginal exposure", "Imaginal exposure (experimental paradigm) + placebo",
        "Imaginal exposure + imagery rescripting", "Implosive Therapy",
        "Virtual reality exposure therapy (VRET)", "Written Exposure Therapy (WET)",
        "Written narrative exposure", "Imagery Rehearsal Training (IRT)",
        "Cognitive Behavioral Conjoint Therapy (CBCT)", "Impact of Killing (IOK)",
        "Trauma-focused cognitive behavioral therapy (TF-CBT)",
        "Cognitive Processing Therapy (CPT)", "Cognitive Therapy for PTSD (CT-PTSD)"
      ) ~ "TF-CBT",

      `Standardized Treatment Name` %in% c(
        "Brief Eclectic Psychotherapy (BEP)", "Modified Lifespan Integration",
        "Multimodular motion-assisted memory desensitization and reconsolidation (3MDR)",
        "Observed and Experiential Integration - Switching component",
        "Reconsolidation of Traumatic Memories (RTM)", "Trauma counseling",
        "Accelerated Resolution Therapy (ART)"
      ) ~ "TF-Other",

      `Standardized Treatment Name` %in% c(
        "Behavioral activation (BA)",
        "Cognitive behavioral therapy (CBT) without a trauma focus",
        "Cognitive Behavioral Therapy for Insomnia (CBT-I)",
        "Cognitive restructuring and imagery modification", "Metacognitive Therapy",
        "Stress Inoculation Training (SIT)", "Affect management therapy",
        "Coping skills treatment", "Interpersonal Psychotherapy (IPT)",
        "Memory specificity training", "Non-trauma-focused counseling",
        "Present Centered Therapy (PCT)", "Psychodynamic psychotherapy",
        "PTSD psychotherapy", "Resilience-oriented treatment",
        "Skills Training in Affect and Interpersonal Regulation (STAIR)",
        "Supportive counseling", "Dialectical Behavior Therapy"
      ) ~ "Non-trauma-focused",

      `Standardized Treatment Name` %in% c(
        "Cognitive behavioral therapy (CBT) + Eye Desensitization and Reprocessing (EMDR)",
        "Cognitive Behavioral Therapy for Insomnia (CBT-I) + Imagery Rehearsal Training (IRT)",
        "Dialectical Behavior Therapy (DBT) + trauma-focused cognitive behavioral therapy (TF-CBT)",
        "Eye Movement Desensitization and Reprocessing (EMDR) + cognitive behavioral therapy (CBT) without a trauma focus",
        "Eye Movement Desensitization and Reprocessing (EMDR) + progressive muscle relaxation (PMR) + group psychoeducation",
        "Narrative Exposure Therapy (NET) + Interpersonal Psychotherapy (IPT)",
        "Prolonged Exposure (PE) + Stress Inoculation Training (SIT)",
        "Skills Training in Affect and Interpersonal Regulation (STAIR) plus modified Prolonged Exposure (PE)"
      ) ~ "Mixed psychotherapy",

      TRUE ~ "Other"
    )
  )
```

```{r identify-common-other}
other_top <- PTSD_military %>%
  filter(category == "Other") %>%
  count(`Treatment Name`, sort = TRUE)

other_top
```

I tried to figure out ways to recode the rest of these into my exisitng categories, but I just don't think its a good use of my time, nor will it actually give me the information I'm looking for. So instead, I want to focus on how PTSD treatment research in military/veteran populations changed over time. To do so, I will use the study publication year variable. 

```{r add-year-var}
PTSD_military <- PTSD_military %>%
  mutate(year = as.numeric(`Study Publication Year`)) %>%
  filter(!is.na(year))
```

```{r count-studies-per-yr-per-category}
year_counts <- PTSD_military %>%
  count(year, category)
```

```{r plot}
ggplot(year_counts,
       aes(x = year, y = n, color = category)) +
  geom_line() +
  labs(
    title = "Trends in PTSD Treatment Research Over Time",
    subtitle = "Military and Veteran Studies",
    x = "Publication Year",
    y = "Number of studies",
    color = NULL
  ) +
  theme_minimal()
```

This visualization shows a substantial increase in PTSD treatment research in military and veteran populations beginning around 2010. However, the majority of studies fall into the "other" category. 
